---
title: "Article"
author: "Philipp Baumann"
date: "2019-12-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Albert Einstein would probably not feel the necessity for simplerspec, as he follows his quote "Everything should be made as simple as possible, but not simpler".

# Prepare the R environment

To reproduce the analysis in this hands-on, I would advise two main options:

1. Manual installation of R packages
2. Installing exact package versions and sources using the renv package and the snapshot file `renv.lock`

## Manual installation

To install and attach all required R packages used in this article, you can run the following  lines:

```{r}
pkgs <- c("here", "simplerspec", "tidyverse", "data.table")
new_pkgs <- pkgs[!(pkgs %in% installed.packages()[, "Package"])]
# Install only new packages
if (length(new_pkgs)) {
  if ("remotes" %in% new_pkgs) install.packages("remotes")
  if ("simplerspec" %in% new_pkgs) {
    remotes::install_github("philipp-baumann/simplerspec")}
  install.packages(new_pkgs)
}
suppressPackageStartupMessages(
  purrr::walk(pkgs, library, character.only = TRUE, quietly = TRUE)
)
```

## Restore the project with renv

First, clone this repository to your local computer. Then install renv and restore R packages based on the `renv.lock` file.

# Outro

```{r}
devtools::session_info()
```


